---
title: "R Basic Data Types"
author: "Sivan Margalit"
date: "`r Sys.Date()`"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Basic Data Types
In this tutorial we focus on 4 basic data types (or classes), used by R : .
<ul>
<li> numeric </li>
<li> character </li>
<li> logical </li>
<li> dateTime </li>
</ul>


 give example 3 first data types: 
 check your example with the function class(...) (write your example between the Parenthesis)
```{r data-type-example, exercise=TRUE,  exercise.lines = 7 }
# Example of Numeric value 
class( )

# Example of character value 

# Example of logical value 

```
### numerics
numeric values without decimal point called <b>integers</b> </br>
we can use the functions
<ul>
<li>is.numeric () to check if value is numeric</li>
<li>is.integer() to check if value is integer</li>
</ul>

for example:
```{r check-numeric, echo = TRUE}
x <- 4.5

is.numeric(x)

is.integer(x)
```
now, your turn : assign an <u>integer</u> to variable y and validate

```{r check-integer,  exercise=TRUE}
y <- 
is.numeric(y)

is.integer(y)
```

<div id="check-integer-hint">
**Hint#1:** integers are numbers without decimal points (e.g : 11, -7, 5 )
</div>

### characters
characters values are textual. those values must be surrounded by double quote ("..") or single quote ('...') </br>
<ul>
<li>is.character () check if value type is characters</li>
</ul>

For example : assign to x a character values and validate it 
```{r check-character,  exercise=TRUE}
x <- 

is.character(x)
```


### logical
logical values are TRUE/FALSE (called also booleans). </br>
<ul>
<li>is.logical() check if value is type logic' i.e. TRUE/FALSE</li>
</ul>
for example: is.numeric() function returns a logical value. </br>
another exampleL  assign to z a logical expression (4<3). add validation if it indeed logical
```{r check-logical, exercise=TRUE }
z <- 4<3
# add here a validation if x is logical

```
```{r check-logical-solution }
z <- 4<3
# add here a validation if x is logical

is.logival(z)  #validate with is.logical() 

class(z) #another validation: check z class ()
```
### Date and dateTime
<ul>
<li>
<b>Date</b> values should be convert from text, use the function as.Date()
```{r convert-date, echo=TRUE }
dd <- as.Date("2022-01-01")
dd
```
In the example above a variable dd assigned to Date value</br>
even though it displayed as a text (character), it is actually a Date.</br>
lets validate it
```{r check-date, exercise=TRUE }
dd <- as.Date("2022-01-01")
# check here if dd is a character 

# check here the class of dd  

```
```{r check-date-solution}
dd <- as.Date("2022-01-01")
# check here if dd is a character 
is.character(dd)
# check here the class of dd  
class(dd)

```
</li>

<li>
<b>POSIXct</b> is special class contains date and time together
```{r convert-dateTime, echo=TRUE }
dt <- as.POSIXct("2022-01-01 10:15")
dt
class(dt)
```
It is easy to extract date from POSIXct, using as.Date()
```{r dateTime-to-Date, echo=TRUE}
dt <- as.POSIXct("2022-01-01 10:15")
dt
as.Date(dt)
```
</li>


## Check yourself

```{r question-character-1, echo=FALSE}
question("which of the following variable is character?",
  answer("A<-'8'", correct = TRUE),
  answer("B<-8"),
  answer("C<-'eight'", correct = TRUE),
  answer("D<-TRUE"),
  allow_retry = TRUE,
  random_answer_order = TRUE,
  try_again_button = "Try Again"
)
```

```{r question-character-2, echo=FALSE}
i <- sample(1:3, 1)
values.lst<-list("'3'"="3",
                 "3"=3,
                 "'three'"="three")
xText<-names(values.lst)[i]
xValue<-values.lst[[i]]

questionText <-sprintf("given x = %s. what expected result of the command is.character(x) :", xText)
question(questionText,
  answer(sprintf("%s", ifelse(!is.character(xValue),"TRUE", "FALSE")), correct = FALSE),
  answer(sprintf("%s", ifelse(is.character(xValue),"TRUE", "FALSE")), correct = TRUE),
  answer(xValue, correct = FALSE),
  answer("error message", correct = FALSE),
  allow_retry = TRUE,
  random_answer_order = TRUE,
  try_again_button = "Try Again"
)
```


```{r question-character-3, echo=FALSE}

question('given y = "2020-12-17". what expected result of the command class(y) ',
  answer("Date", correct = TRUE),
  answer("FALSE"),
  answer("33.4"),
  answer("error message"),
  allow_retry = TRUE,
  random_answer_order = TRUE,
  try_again_button = "Try Again"
)
```
