---
title: "R Data Sructures"
author: "Sivan Margalit"
date: "`r Sys.Date()`"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

This is very quick practice on vectors, matrix, lists and data.frame to remind the basic usage

## Vectors, Matrics and Arrays

### Vectors 
one dimension array of items , **all** of same class </br>

given vector vec_1
```{r define-vector, echo=FALSE}
vec1 <-sample(-100:100, 20, replace = TRUE)
```
in the next code box :</br>
<ul>
<li> check which data type the elements in the vector</li>
<li> the size of the vector </li>
<li> summarize the vector </li>
<li> print the values in the 1st, 3rd, 6th places</li>
</ul>

```{r vector-functions, exercise=TRUE, exercise.lines = 9 }
# 1. check which data type the elements in the vector (use function class)

# 2. check the size of the vector (use length function)

# 3. summarise vector values distributon (use summary function)

# 4. print the values in the 1st, 3rd, 6th places (use operator [ ])

```
```{r vector-functions-solution}
# 1. check which data type the elements in the vector (use function class)
class(vec_1)
# 2. check the size of the vector (use length function)
length(vec_1)
# 3. summaries vector values distributon (use summary function)
summary(vec_1)
# 4. print the values in the 1st, 3rd, 6th places (use operator [ ])
vec_1[c(1, 3, 6)]
```
 icalture")    # vector of textual values

c((1:5) < 3)    # vector of logical values

```
### matrix
2 dimensions array of items. **all** of same class 

```{r create-matrix, echo=TRUE}
# matrix(...) function generate  vector

matrix(1:27, nrow=3) # matrix of integers, in 3 rows

matrix(letters, ncol=3)    # matrix of characters (textuals)

```
### array
array of items, in any dimensions. **all** of same class 

```{r create-array, echo=TRUE}
# array(...) function generate  array

array(1:28, dim=c(7,2,2)) # array of integers: 7 rows, 2 columns, 2 blocks

array(1:28, dim=c(2,7,2)) # same values but different structur: 2 rows, 7 columns, 2 blocks

```
### practices


#### <u>numerics</u>
numeric values without decimal point called <b>integers</b> </br>
we can use the functions
<ul>
<li>is.numeric () to check if value is numeric</li>
<li>is.integer() to check if value is integer</li>
</ul>

for example:
```{r check-numeric, echo = TRUE}
x <- 4.5

is.numeric(x)

is.integer(x)
```
now, your turn : assign an <u>integer</u> to variable y and validate

```{r check-integer, exercise=TRUE}
y <- 
is.numeric(y)

is.integer(y)
```

<div id="check-integer-hint">
**Hint#1:** integers are numbers without decimal points (e.g : 11, -7, 5 )
</div>

#### <u>characters</u>
characters values are textual. those values must be surrounded by double quote ("..") or single quote ('...') </br>
<ul>
<li>is.character () check if value type is characters</li>
</ul>

For example : assign to x a character values and validate it 
```{r check-character,  exercise=TRUE}
x <- 

is.character(x)
```


#### <u>logical</u>
logical values are TRUE/FALSE (called also booleans). </br>
for example: is.numeric() function returns a logical value
<ul>
<li>is.logical() check if value is type logic' i.e. TRUE/FALSE</li>
</ul>

For example : assign to z a logical values and validate it 
```{r check-logical, exercise=TRUE }
x <- 4<3
# add here a validation if x is logical

```

### Check yourself

```{r question-character-1, echo=FALSE}
question("wich of the following variable is character?",
  answer("A<-'8'", correct = TRUE),
  answer("B<-8"),
  answer("C<-'eight'", correct = TRUE),
  answer("D<-TRUE"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

```{r question-character-2, echo=FALSE}
i <- sample(1:3, 1)
values.lst<-list("'3'"="3",
                 "3"=3,
                 "'three'"="tree")
xText<-names(values.lst)[i]
xValue<-values.lst[[i]]

questionText <-sprintf("given x = %s. what expected result of the command is.character(x) :", xText)
question(questionText,
  answer(sprintf("%s", ifelse(!is.character(xValue),"TRUE", "FALSE")), correct = FALSE),
  answer(sprintf("%s", ifelse(is.character(xValue),"TRUE", "FALSE")), correct = TRUE),
  answer(xValue, correct = FALSE),
  answer("error message", correct = FALSE),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```


```{r question-character-3, echo=FALSE}
z <- 33.4
is.numeric(z)

question("given x = 33.4. what expected result of the command is.numeric(z) ",
  answer("TRUE", correct = TRUE),
  answer("FALSE"),
  answer("33.4"),
  answer("error message"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```
## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaluated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

